---
title: "DiaOutline and EFA statistical analysis"
author: "Soren Pedersen"
date: "18/10/2021"
output: html_document
---

(1.1) First install DiaOutline from https://github.com/wishkerman/DiaOutline where the three files (DiaOutline.exe / DiaOutline101.7z.001 / DiaOutline101.7z.002) are downloaded. Download the .exe file last and make sure that the .001 and .002 files are run using the .exe file as operator. Read the readme file.

(1.2) Install MATLAB Runtime (R2017b) from http://www.mathworks.com/products/compiler/mcr/index.html and install the necessary updates. Make a profile on the site. Run as admin. 

(1.3) Install these packages for statistical analysis (Wishkerman et al 2018)

```{r, echo=TRUE}
install.packages("MASS")
install.packages("ggplot2")
install.packages("GGally")
install.packages("doBy")
install.packages("data.table")
install.packages("plyr")
install.packages("grid")
install.packages("gridExtra")
install.packages("Momocs")
```
(1.4) and load
```{r, echo=TRUE}
library(MASS)
library(ggplot2)
library(GGally)
library(doBy)
library(data.table)
library(plyr)
library(grid)
library(gridExtra)
library(Momocs)
```

